---
title: "Assignment 1"
author: "Christina Lin"
date: "9/17/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1. Retrieving data and checking for missing values

Reading 2004 and 2019 Data
```{r}
library(data.table)
library(tidyverse)

if (!file.exists("../2004_data"))
  download.file(
    url = "https://www3.epa.gov/cgi-bin/broker?_service=data&_server=134.67.99.91&_port=4090&_sessionid=7BDG2Jm/R52&_PROGRAM=dataprog.ad_viz_plotval_getdata.sas",
    destfile = "2004_data",
    method   = "libcurl",
    timeout  = 60
    )
pm2004 <- data.table::fread("2004_data")


if (!file.exists("../2019_data"))
  download.file(
    url = "https://www3.epa.gov/cgi-bin/broker?_service=data&_server=134.67.99.91&_port=4077&_sessionid=xPDTDJm/R52&_PROGRAM=dataprog.ad_viz_plotval_getdata.sas",
    destfile = "2019_data",
    method   = "libcurl",
    timeout  = 60
    )
pm2019 <- data.table::fread("2019_data")

```

Checking dimensions, headers, footers, variables names and variable types

```{r}
# For 2004 Dataset
dim(pm2004)
head(pm2004)
tail(pm2004)
str(pm2004)


# For 2019 Dataset
dim(pm2019)
head(pm2019)
tail(pm2019)
str(pm2019)

```

Checking for issues in key variables
```{r}
summary(pm2004$`Daily Mean PM2.5 Concentration`)
summary(pm2004$SITE_LATITUDE)
summary(pm2004$SITE_LONGITUDE)
mean(is.na(pm2004$`Daily Mean PM2.5 Concentration`))
mean(is.na(pm2004$SITE_LATITUDE))
mean(is.na(pm2004$SITE_LONGITUDE))


summary(pm2019$`Daily Mean PM2.5 Concentration`)
summary(pm2019$SITE_LATITUDE)
summary(pm2019$SITE_LONGITUDE)
mean(is.na(pm2019$`Daily Mean PM2.5 Concentration`))
mean(is.na(pm2019$SITE_LATITUDE))
mean(is.na(pm2019$SITE_LONGITUDE))

```



In 2004, there were 19233 observations total. In 2019, there were 53086 observations total. There are no missing values (i.e. "NA"s) for daily mean 2.5 concentration, latitude, and longitude for both the 2004 and 2019 datasets. Upon examining the minimum and maximum values for these variables, there does not appear to be any "9999" in place of missing values either. Thus, the datasets looks ready for analysis. 



# Step 2: Combining the two data sets into one data frame

```{r}
library(dplyr)
pm <- rbind(pm2004, pm2019)

#Adding year as a new column and simplifying names of key variables 
pm <- mutate(pm, year = factor(rep(c(2004, 2019), c(nrow(pm2004), nrow(pm2019))))) %>%
  rename(dPM ="Daily Mean PM2.5 Concentration") %>%
  rename(lat = SITE_LATITUDE) %>%
  rename(lon = SITE_LONGITUDE) %>%
  rename(site = "Site ID") %>%
  rename(site_name = "Site Name")
str(pm)

```



# Step 3: Create a basic map in leaflet() that shows the locations of the sites

```{r}

library(leaflet)

pal <- colorFactor(palette=c('red','blue'), domain=pm$year)

leaflet(pm) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
    color = ~pal(year),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend('bottomleft', pal=pal, values=pm$year,
          title='Year', opacity=1)

```


Observations: The sites in 2004 were mainly at the north and south ends of California while in 2019, the sites are more evenly scattered around the state. 



# Step 4: Check for any missing or implausible values of PM in the combined dataset

```{r}
summary(pm)
```


Observations: There are PM2.5 values that are negative, which does not seem plausible since these values represent a concentration. 


```{r}
#Filtering out the data points that have negative PM2.5 values
pm2 <- pm[dPM < 0]
summary(pm2)

pal2 <- colorFactor(palette=c('red','blue'), domain=pm2$year)

leaflet(pm2) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
    color = ~pal2(year),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend('bottomleft', pal=pal2, values=pm2$year,
          title='Year', opacity=1)

```


Observations: There is 1 negative PM2.5 value recorded in 2004 and 282 of such values in 2019. By mapping these data points, it appears that the negative value recorded in 2004 is located in the middle of the state. The negative values recorded in 2019 are rather evenly spread out throughout the state. 



# Step 5: Explore the main question of interest at three different spatial levels

Comparing PM2.5 levels between 2004 and 2019 at the state level

```{r}
ggplot(data = pm, mapping = aes(x = year, y = dPM)) +
  geom_boxplot()

ggplot(data = pm, mapping = aes(x = dPM)) +
  geom_histogram(bins=20) +
  facet_grid(~year)

summary(pm[year==2004 & dPM >= 0,dPM])
summary(pm[year==2019 & dPM >=0 ,dPM])
```


Observations: The average and media PM2.5 levels appear lower in 2019 than 2004 and the maximum PM2.5 level measured is lower in 2019. This is the case when negative values are included but also when negative values are excluded from the statistics summary. 



Comparing PM2.5 levels between 2004 and 2019 at the county level

```{r}
#Averaging daily PM2.5 values, grouped by year and county code 
pm_county_avg <- group_by(pm, year, COUNTY_CODE) %>%
  summarize(dPM = mean(dPM))

head(pm_county_avg)
tail(pm_county_avg)

#Making line graph to show change in PM2.5 values in counties between the two years
ggplot(data = pm_county_avg, mapping = aes(x = year,y = dPM, color = as.factor(COUNTY_CODE))) +
  geom_point() +
  geom_line(aes(group= as.factor(COUNTY_CODE)))
```


Observations: In most counties, the average PM2.5 values have decreased from 2004 to 2019. There are a few counties that saw increased average PM2.5 values, but these increases are not very big. 



Comparing PM2.5 levels between 2004 and 2019 at Los Angeles site

```{r}
pm_la <- filter(pm, site_name == "Los Angeles-North Main Street")

ggplot(data = pm_la, mapping = aes(x = year, y = dPM)) +
  geom_boxplot()

ggplot(data = pm_la, mapping = aes(x = dPM)) +
  geom_histogram(bins=20) +
  facet_grid(~year)

```


Observations: At the site in LA, the average PM2.5 values are lower in 2019 compared to 2004. The histogram also shows that there are more observations at lower PM2.5 values in 2019 compared to 2004. 



Conclusion: Daily PM 2.5 values in California appear to have decreased from 2004 to 2019.
