---
title: "Assignment 4"
author: "Christina Lin"
date: "11/16/2021"
output:
  html_document:
    html_preview: false
  github_document: default
  word_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Part 1: HPC

## Problem 1

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  t(apply(X = mat,MARGIN = 1, FUN = cumsum))
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)

```


## Problem 2

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132

```


```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Now re-writing using parLapply:

```{r}
library(parallel)

system.time({
  cl <- makePSOCKcluster(4L)
  clusterSetRNGStream(cl,1231)
  clusterExport(cl,"sim_pi")
  ans <- unlist(parLapply(cl, rep(4000,10000),sim_pi))
  print(mean(ans))
  stopCluster(cl)
})
```


# Part 2: SQL

```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```


## Question 1
How many movies aew there avaliable in each rating catagory.
```{sql connection=con}
SELECT rating, COUNT (*) AS count
FROM film
GROUP BY rating
```


## Question 2
What is the average replacement cost and rental rate for each rating category.
```{sql connection=con}
SELECT 
  AVG(replacement_cost) AS avg_replacement_cost,
  AVG(rental_rate) AS avg_rental_rate,
  rating
FROM film
GROUP BY rating

```

## Question 3
Use table film_category together with film to find the how many films there are with each category ID

```{sql connection=con}
SELECT COUNT(*) AS "N_Films", category_id
FROM film
INNER JOIN film_category
ON film.film_id = film_category.film_id
GROUP BY category_id
```


## Question 4
Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql connection=con}

SELECT COUNT(*) AS "N_Films", name
FROM film
  INNER JOIN film_category
    ON film.film_id = film_category.film_id
  INNER JOIN category
    ON film_category.category_id = category.category_id
GROUP BY name
ORDER BY N_Films DESC

```


