---
title: "Assignment 2"
author: "Christina Lin"
date: "10/1/2021"
output: 
  github_document:
    html_preview: false
  html_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(data.table)
library(tidyverse)

if (!file.exists("individual.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv",
    destfile = "individual.csv",
    method   = "libcurl",
    timeout  = 60
    )
individual <- fread("individual.csv")


if (!file.exists("regional.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv",
    destfile = "regional.csv",
    method   = "libcurl",
    timeout  = 60
    )
regional <- fread("regional.csv")

```

# Data Wrangling

## Question 1: Merging the data sets

```{r}
merged <- merge(
  x = individual,
  y = regional,
  all.x = TRUE, all.y = FALSE,
  by.x = "townname", by.y = "townname"
)

```


Checking dimensions to see if merge was successful
```{r}
dim(individual)
dim(regional)
dim(merged)
```

There are `r nrow(individual)` rows in the individual data set and `r nrow(regional)` rows in the regional data set. After merging, the new data set has `r nrow(merged)` rows, indicating a successful merge. 




Imputing missing values of important variables with averages within "male" and "hispanic"

```{r}
bmi_avg <- mean(merged[male==1 & hispanic==1, bmi], na.rm = TRUE)
fev_avg <- mean(merged[male==1 & hispanic==1, fev], na.rm = TRUE)
asthma_avg <- mean(merged[male==1 & hispanic==1, asthma], na.rm = TRUE)
gasstove_avg <- mean(merged[male==1 & hispanic==1, gasstove], na.rm = TRUE)
smoke_avg <- mean(merged[male==1 & hispanic==1, smoke], na.rm = TRUE)


merged[is.na(bmi), bmi := bmi_avg]
merged[is.na(fev), fev := fev_avg]
merged[is.na(asthma), asthma := asthma_avg]
merged[is.na(gasstove), gasstove := gasstove_avg]
merged[is.na(smoke), smoke := smoke_avg]

head(merged)

```


## Question 2: Creating BMI categorical variable and summary table

```{r}
merged[bmi < 14, obesity_level := "underweight"]
merged[bmi >= 14 & bmi < 21, obesity_level := "normal"]
merged[bmi >= 21 & bmi < 24, obesity_level := "overweight"]
merged[bmi > 24, obesity_level := "obese"]

tab <- merged[, .(
  N_Entries = .N,
  min_BMI = min(bmi),
  max_BMI = max(bmi)
), by = obesity_level]

knitr::kable(tab)
```


## Question 3: Creating Smoke and Gas Exposure categorical variable 

```{r}
merged[smoke == 0 & gasstove == 0, smoke_gas_exposure := "none"]
merged[smoke == 0 & gasstove == 1, smoke_gas_exposure := "only gas"]
merged[smoke == 1 & gasstove == 0, smoke_gas_exposure := "only smoke"]
merged[smoke == 1 & gasstove == 1, smoke_gas_exposure := "both"]

```



## Question 4: Summary of FEV and Asthma indicator by town, sex, obesity level, and exposure

```{r}
tab_town <- merged[, .(
  avg_fev = mean(fev),
  sd_fev = sd(fev),
  asthma_proportion = mean(asthma),
  asthma_sd = sd(asthma)), by = townname]
knitr::kable(tab_town)


tab_sex <- merged[, .(
  avg_fev = mean(fev),
  sd_fev = sd(fev),
  asthma_proportion = mean(asthma),
  asthma_sd = sd(asthma)), by = male]
knitr::kable(tab_sex)

tab_obesity <- merged[, .(
  avg_fev = mean(fev),
  sd_fev = sd(fev),
  asthma_proportion = mean(asthma),
  asthma_sd = sd(asthma)), by = obesity_level]
knitr::kable(tab_obesity)

tab_exposure <- merged[, .(
  avg_fev = mean(fev),
  sd_fev = sd(fev),
  asthma_proportion = mean(asthma),
  asthma_sd = sd(asthma)), by = smoke_gas_exposure]
knitr::kable(tab_exposure)
```


# Looking at the Data

## 1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r}
ggplot(merged, mapping = aes(x = bmi, y = fev)) +
  geom_point()+
  geom_smooth(method = lm, se = FALSE)+
  facet_wrap(~townname)
```

Observations: In most of the towns, there appears to be a positive correlation between FEV and BMI. Towns that show little association are Riverside and Alpine. 



## 2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r}
color_pal <- c("pink","violet", "darkviolet", "hot pink")

ggplot(merged, mapping = aes(x=fev, fill = obesity_level)) +
  geom_histogram(col = "black", bins = 20) +
  scale_fill_manual(values = color_pal)


ggplot(merged, mapping = aes(x=fev, fill = smoke_gas_exposure)) +
  geom_histogram(col = "black", bins = 20) +
  scale_fill_brewer(palette = "Greens")

```

Observations: From the FEV-BMI histogram, it appears that there are more observations of higher FEV in the overweight and obese categories. There are also more observations of lower FEV in underweight and normal categories. This supports the trend seen in the scatterplots above with positive correlation between FEV and BMI. From the FEV-Smoke/Gas Exposure histogram, it appears that there are more observations of lower FEV with gas exposure. The trend between FEV and smoke exposure is difficult to tell from the histogram.



## 3. Bar chart of BMI by smoke/gas exposure.

```{r}
ggplot(merged, mapping = aes(x = smoke_gas_exposure, y = bmi)) +
  geom_bar(stat= "summary")

```

Observations: This graph indicates little difference in bmi between the different smoke and gas exposure groups.  



## 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

```{r}
ggplot(merged, mapping = aes(x = smoke_gas_exposure, y = fev)) +
  geom_boxplot()

ggplot(merged, mapping = aes(x = smoke_gas_exposure, y = bmi)) +
  geom_boxplot()

```

Observations: The bar plots indicate that the median and quartile values are rather similar between the different smoke and gas exposure groups in terms of fev and bmi. The median bmi is slightly higher in the only smoke exposure group. 



## 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r eval = knitr::is_html_output(excludes = "gfm")}
library(leaflet)

pal <- colorNumeric(palette="viridis", domain=merged$pm25_mass)

leaflet(merged) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
    color = ~pal(pm25_mass),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend('bottomleft', pal=pal, values=merged$pm25_mass,
          title='PM2.5 Mass', opacity=1)


```

Observations: The highest PM2.5 mass is found in communities around Riverside, Long Beach, and San Bernardino. 




## 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}
ggplot(merged, mapping = aes(x = pm25_mass, y = fev)) +
  geom_point()+
  geom_smooth(method = lm, se = FALSE)+
  geom_jitter()
```

Observations: There may be a trend towards a negative association between fev and pm25_mass, but it is not obvious and is likely not a significant association. 



## Conclusions
There is a positive association between BMI and FEV. However, there are no obvious associations between FEV and smoke/gas exposure as well as FEV and PM2.5 exposure. 
