---
title: "Assignment 2"
author: "Christina Lin"
date: "10/1/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(data.table)
library(tidyverse)

if (!file.exists("individual.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv",
    destfile = "individual.csv",
    method   = "libcurl",
    timeout  = 60
    )
individual <- fread("individual.csv")


if (!file.exists("regional.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv",
    destfile = "regional.csv",
    method   = "libcurl",
    timeout  = 60
    )
regional <- fread("regional.csv")

```

Merging the data sets

```{r}
merged <- merge(
  x = individual,
  y = regional,
  all.x = TRUE, all.y = FALSE,
  by.x = "townname", by.y = "townname"
)

```


Checking dimensions to see if merge was successful
```{r}
dim(individual)
dim(regional)
dim(merged)
```

There are `r nrow(individual)` rows in the individual data set and `r nrow(regional)` rows in the regional data set. After merging, the new data set has `r nrow(merged)` rows, indicating a successful merge. 




Imputing missing values of important variables with averages within "male" and "hispanic"

```{r}
bmi_avg <- mean(merged[male==1 & hispanic==1, bmi], na.rm = TRUE)
fev_avg <- mean(merged[male==1 & hispanic==1, fev], na.rm = TRUE)


merged[is.na(bmi), bmi := bmi_avg]
merged[is.na(fev), fev := fev_avg]

head(merged)

```



Creating BMI categorical variable

```{r}
merged[bmi < 14, obesity_level := "underweight"]
merged[bmi >= 14 & bmi < 21, obesity_level := "normal"]
merged[bmi >= 21 & bmi < 24, obesity_level := "overweight"]
merged[bmi > 24, obesity_level := "obese"]

tab <- merged[, .(
  N_Entries = .N,
  min_BMI = min(bmi),
  max_BMI = max(bmi)
), by = obesity_level]

knitr::kable(tab)
```



